---
title: "GSoC2021: tests"
author: "Haris Zafeiropoulos"
date: "4/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## From DNA sequences to metabolic interactions: building a pipeline to extract key metabolic processes

Based on previous work, I propose a Google Summer of Code project in the framework of the [GeomScale group]()
that is not listed in their [table of proposed coding projects](https://github.com/GeomScale/gsoc2021/wiki/table-of-proposed-coding-projects).

However, as it has a lot in common with the 
[Inferring microbial interactions](https://github.com/GeomScale/gsoc2021/wiki/Inferring-microbial-interactions) project,
this is an implementation of the tests that correspond to the later. 



### Easy 

***Question:***
*Compile and run `volesti`. Use the R extension to visualize sampling in a polytope*

After cloning the [`volume_approximation`](https://github.com/GeomScale/volume_approximation/) repository and 
checking out to the `branch` of the [`SoCG21`](https://github.com/GeomScale/volume_approximation/tree/socg21)

I installed the [dependencies](https://github.com/GeomScale/volume_approximation/tree/socg21) described there 
to install this `volesti` version, including some extra, metabolic-networks-oriented functions.

To sample on a polytope, I got an example of a cube polytope from the `R-proj/inst/extdata` directory of the repo.
You may find this cube [here](https://github.com/GeomScale/volume_approximation/blob/develop//inst/extdata/cube_10.ext)

```{r cars}
library(volesti)
library(ggplot2)
setwd("/home/haris/Documents/coding/github_repos/volume_approximation/R-proj/")
polytope = file_to_polytope("examples/cube10.ine")
```

As I chose an `.ine` file, I got a H-polytope.
I then checked the dimension of the cube polytope. 

```{r dimension}
polytope$dimension
```

Then I sampled on the cube polytope with various random walks supported by the `volesti` library, 
asking for 10,000 points in every case. 

```{r sampling_on_cube}
# 'CDHR' for Coordinate Directions Hit-and-Run
points_cdhr =  sample_points(P = polytope, n = 10000, 
                             distribution = list("density" = "uniform"),  random_walk = list("walk"="CDHR"))
CDHR = data.frame(x=points_cdhr[1,], y=points_cdhr[2,])

# 'RDHR' for Random Directions Hit-and-Run
points_rdhr =  sample_points(P = polytope, n = 10000, 
                             distribution = list("density" = "uniform"),  random_walk = list("walk"="RDHR"))
RDHR = data.frame(x=points_rdhr[1,], y=points_rdhr[2,])

# BiW' for Billiard walk
points_biw =  sample_points(P = polytope, n = 10000, 
                            distribution = list("density" = "uniform"),  random_walk = list("walk"="BiW"))
BiW = data.frame(x=points_biw[1,], y=points_biw[2,])

# BaW' for Ball Walk
points_baw =  sample_points(P = polytope, n = 10000, 
                            distribution = list("density" = "uniform"),  random_walk = list("walk"="BaW"))
BaW = data.frame(x=points_baw[1,], y=points_baw[2,])

# 'dikin' for dikin walk
points_dikin =  sample_points(P = polytope, n = 10000, 
                            distribution = list("density" = "uniform"),  random_walk = list("walk"="dikin"))
dikin = data.frame(x=points_dikin[1,], y=points_dikin[2,])

# 'vaidya' for vaidya walk, 
points_vaidya =  sample_points(P = polytope, n = 10000, 
                            distribution = list("density" = "uniform"),  random_walk = list("walk"="vaidya"))
vaidya = data.frame(x=points_vaidya[1,], y=points_vaidya[2,])

# 'john' for john walk
points_john =  sample_points(P = polytope, n = 10000, 
                            distribution = list("density" = "uniform"),  random_walk = list("walk"="john"))
john = data.frame(x=points_john[1,], y=points_john[2,])

# 'BCDHR' boundary sampling by keeping the extreme points of CDHR 
points_BCDHR =  sample_points(P = polytope, n = 10000, 
                            distribution = list("density" = "uniform"),  random_walk = list("walk"="BCDHR"))
BCDHR = data.frame(x=points_BCDHR[1,], y=points_BCDHR[2,])

```


And then make a plot to see the distribution of each of these walks! 

```{r plot_samples, echo = FALSE}
walks = rbind(CDHR, RDHR, BiW, BaW, dikin, vaidya, john, BCDHR)
walks$random_walk = c(rep("CDHR",10000),rep("RDHR",10000), rep("BiW", 10000), rep("BaW",10000), rep("dikin",10000), rep("vaidya",10000), rep("john",10000), rep("BCDHR", 10000))

ggplot(walks, aes(x=x, y=y, group=random_walk, col=random_walk)) +
  geom_point() + labs(title = "uniform sampling on cube_10.ine file / H-polytope")
```


### Medium

***Question:***
*Use the R extension to sample on the flux space of the metabolic network of E.coli*

To sample on the flux space of *E.coli*'s metabolic network, we first need to get such a network! :-) 
Metabolic networks can be usually found as a `.json` or/and a `.xml` or/and a `.mat` file. 
We will stick with the `.mat` format. 

```{r e_coli_met_net}
data_url = "http://bigg.ucsd.edu/static/models/e_coli_core.mat"
e_coli_core = "/home/haris/Documents/projects/GeomScale/gsoc2021/e_coli_core.mat"
download.file(data_url, e_coli_core)
```

Now, a single `volesti` function will sample on the flux space of the metabolic network. 

```{r read_mat}

## request effectiveness n = 1000 and sample steady states
result_list = generate_steady_states(e_coli_core, n = 1000, Recon2D_v04 = FALSE, Recon3D_301 = FALSE)

```










